---
title: "Proj2_EDA_JuliaLayne"
author: "Julia Layne"
date: "7/12/2020"
output: word_document
---
# Data set
https://archive.ics.uci.edu/ml/datasets/bank%20marketing
# Paper Referenced - that uses the dataset and conclusions
http://media.salford-systems.com/video/tutorial/2015/targeted_marketing.pdf

## Response
21 - y - has the client subscribed a term deposit? (binary: 'yes','no')

# Questions about dataset
I'm not sure what the following variables mean in the context of the response:

11 - duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.

# other attributes:
12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
14 - previous: number of contacts performed before this campaign and for this client (numeric)
15 - poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')

# social and economic context attributes
16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
17 - cons.price.idx: consumer price index - monthly indicator (numeric)
18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric)
19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
20 - nr.employed: number of employees - quarterly indicator (numeric)



```{r}
library(naniar)
library(gplots)
library(ggplot2)
library(dplyr)

#Bank
setwd('~/Documents/GitHub/Project2_Applied_Stats/Data/')
bank <- read.csv2("bank-additional-full.csv",header = TRUE)
summary(bank)
#pdays may be a irrelevant data point
#Response is heavily weighted on No's 

# ?? WOULD WE EVEN NEED TO SCALE ANY OF THESE CATEGORICAL VARIABLES? Q for Turner

#No variables missing?
gg_miss_var(bank)
summary(bank)
bank
```

```{r}

#Clean up whitespaces

#Unknowns
clean_bank<-bank

#1 - age (numeric)
# okay
#2 - job : type of job (categorical: 'admin.','blue-collar','entrepreneur','housemaid','management','retired','self-employed','services','student','technician','unemployed','unknown')
clean_bank$job <- factor(clean_bank$job)
#3 - marital : marital status (categorical: 'divorced','married','single','unknown'; note: 'divorced' means divorced or widowed)
clean_bank$marital <- factor(clean_bank$marital)
#4 - education (categorical: 'basic.4y','basic.6y','basic.9y','high.school','illiterate','professional.course','university.degree','unknown')
clean_bank$education <- factor(clean_bank$education)
#5 - default: has credit in default? (categorical: 'no','yes','unknown')
clean_bank$default <- factor(clean_bank$default)
#6 - housing: has housing loan? (categorical: 'no','yes','unknown')
clean_bank$housing <- factor(clean_bank$housing)
#7 - loan: has personal loan? (categorical: 'no','yes','unknown')
clean_bank$loan <- factor(clean_bank$loan)


# related with the last contact of the current campaign:
#8 - contact: contact communication type (categorical: 'cellular','telephone') 
clean_bank$contact <- factor(clean_bank$contact)
#9 - month: last contact month of year (categorical: 'jan', 'feb', 'mar', ..., 'nov', 'dec')
clean_bank$month <- factor(clean_bank$month)
#10 - day_of_week: last contact day of the week (categorical: 'mon','tue','wed','thu','fri')
clean_bank$day_of_week <- factor(clean_bank$day_of_week)
#11 - duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.


# other attributes:
#12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
#13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
#14 - previous: number of contacts performed before this campaign and for this client (numeric)
#15 - poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')
# social and economic context attributes
clean_bank$poutcome <- factor(clean_bank$poutcome)
#16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
clean_bank$int.emp.var.rate<-as.integer(clean_bank$emp.var.rate)
clean_bank$emp.var.rate <- factor(clean_bank$emp.var.rate)
#17 - cons.price.idx: consumer price index - monthly indicator (numeric) 
clean_bank$int.cons.price.idx<-as.integer(clean_bank$cons.price.idx)
clean_bank$cons.price.idx <- factor(clean_bank$cons.price.idx)
#18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric) 
clean_bank$int.cons.conf.idx<-as.integer(clean_bank$cons.conf.idx)
clean_bank$cons.conf.idx <- factor(clean_bank$cons.conf.idx)
#19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
clean_bank$int.euribor3m<-as.integer(clean_bank$euribor3m)
clean_bank$euribor3m <- factor(clean_bank$euribor3m)
#20 - nr.employed: number of employees - quarterly indicator (numeric)
clean_bank$int.nr.employed<-as.integer(clean_bank$nr.employed)
clean_bank$nr.employed <- factor(clean_bank$nr.employed)

#Output variable (desired target):
#21 - y - has the client subscribed a term deposit? (binary: 'yes','no')
clean_bank$y <- factor(clean_bank$y)

clean_bank
```

```{r}
int_bank_fields <- c("age","campaign","pdays","previous","int.emp.var.rate","int.cons.price.idx", "int.emp.var.rate","int.cons.price.idx","int.euribor3m","int.nr.employed")

int_bank_fields_y <- c("age","campaign","pdays","previous","int.emp.var.rate","int.cons.price.idx", "int.emp.var.rate","int.cons.price.idx","int.euribor3m","int.nr.employed","y")
clean_bank[int_bank_fields_y]
```

```{r}
bank_factors <- c("job", "marital", "education","default","housing","loan", "contact", "month", "day_of_week","poutcome","emp.var.rate","cons.price.idx","cons.conf.idx","euribor3m","nr.employed","y")

summary(clean_bank[bank_factors])
```


```{r}
cols <- character(nrow(clean_bank))
cols[] <- "black"

cols[clean_bank$y == "no"] <- "blue"
cols[clean_bank$y == "yes"] <- "red"

pairs(clean_bank[c("age","campaign","pdays","previous")],col=cols)
```
```{r}
clean_bank[int_bank_fields]
pairs(clean_bank[int_bank_fields],col=cols)

```
```{r}

cor(subset(clean_bank, select=int_bank_fields)) -> df_cortest 
melt(df_cortest) %>% ggplot(aes(x=Var1, y = Var2, fill = value)) + geom_tile(color = "white")+scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1,1), space = "Lab", name="Pearson\nCorrelation") + theme_minimal()+ theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1))+coord_fixed()


#cor(subset(clean_bank, select=-int_bank_fields)) -> df_cortest 
#melt(df_cortest) %>% ggplot(aes(x=Var1, y = Var2, fill = value)) + geom_tile(color = "white")+scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1,1), space = "Lab", name="Pearson\nCorrelation") + theme_minimal()+ theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1))+coord_fixed()

```


```{r}
pc.bc<-prcomp(clean_bank[int_bank_fields],scale.=TRUE)
pc.bc.scores<-pc.bc$x

#Adding the response column to the PC's data frame
pc.bc.scores<-data.frame(pc.bc.scores)
pc.bc.scores$Diagnosis<-clean_bank[int_bank_fields_y]$y

#Use ggplot2 to plot the first few pc's
library(ggplot2)
ggplot(data = pc.bc.scores, aes(x = PC1, y = PC2)) +
  geom_point(aes(col=Diagnosis), size=1)+
  ggtitle("Yes/No")

ggplot(data = pc.bc.scores, aes(x = PC2, y = PC3)) +
  geom_point(aes(col=Diagnosis), size=1)+
  ggtitle("Yes/No")
```

```{r}
library(Lahman)
library(MASS)
mylda<-lda(y~.,data=clean_bank[int_bank_fields_y])

summary(mylda)
pred<-predict(mylda,newdata=clean_bank[int_bank_fields_y])$class  #Predictions can come in many forms, the class form provides the categorical level of your response.
Truth<-clean_bank[int_bank_fields_y]$y
x<-table(pred,Truth) # Creating a confusion matrix
x
#Missclassification Error
ME<-(x[2,1]+x[1,2])/100
ME
#Calculating overall accuracy
1-ME
```

```{r}
library(caret)
downSample(clean_bank[int_bank_fields_y],clean_bank[int_bank_fields_y]$y)

clean_bank[int_bank_fields]

```







```{r}
mymat<-matrix(c(76,335,105,302),2,2,byrow=T,dimnames=list(c("Placebo","Vitamin C"),c("No","Yes")))
mymat
```


```{r}
field <- "contact"
fct_y <- c(field, "y")
factors_in <- levels(clean_bank[c(field)])

factors_in

tele <- filter(clean_bank, contact == 'telephone')
cell <- filter(clean_bank, contact == 'cellular')

nrow(filter(tele[fct_y], y == 'yes')[c("contact")])


tele_yes <- nrow(filter(tele[fct_y], y == 'yes'))
cell_yes <- nrow(filter(cell[fct_y], y == 'yes'))
tele_no <- nrow(filter(tele[fct_y], y == 'no'))
cell_no <- nrow(filter(cell[fct_y], y == 'no'))

matrx_contact <- c( tele_no, tele_yes, cell_no, cell_yes)
matrx_contact
```


```{r}
library(epitools)
#Another way to format a count matrix
mymat2<-matrix(matrx_contact,2,2,byrow=T)
dimnames(mymat2)<-list("Treatment"=c("Tele","Cell"),"Response"=c("No","Yes"))
mymat2
#Odds Ratio Intervals
oddsratio.wald(mymat2)

#Relative Risk Intervals
riskratio.wald(mymat2)
```  