---
title: "jamie_EDA"
author: "Jamie Vo"
date: "7/12/2020"
output: html_document
---

```{r}
# read in libraries

library(naniar)
library(ggplot2)
library(dplyr)

```

# MARKETING CAMPAIGNS
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATASET 
https://archive.ics.uci.edu/ml/datasets/bank%20marketing

### Variables
Input variables:
#### bank client data:
1 - age (numeric)
2 - job : type of job (categorical: 'admin.','blue-collar','entrepreneur','housemaid','management','retired','self-employed','services','student','technician','unemployed','unknown')
3 - marital : marital status (categorical: 'divorced','married','single','unknown'; note: 'divorced' means divorced or widowed)
4 - education (categorical: 'basic.4y','basic.6y','basic.9y','high.school','illiterate','professional.course','university.degree','unknown')
5 - default: has credit in default? (categorical: 'no','yes','unknown')
6 - housing: has housing loan? (categorical: 'no','yes','unknown')
7 - loan: has personal loan? (categorical: 'no','yes','unknown')
#### related with the last contact of the current campaign:
8 - contact: contact communication type (categorical: 'cellular','telephone')
9 - month: last contact month of year (categorical: 'jan', 'feb', 'mar', ..., 'nov', 'dec')
10 - day_of_week: last contact day of the week (categorical: 'mon','tue','wed','thu','fri')
11 - duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.
#### other attributes:
12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
14 - previous: number of contacts performed before this campaign and for this client (numeric)
15 - poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')
#### social and economic context attributes
16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
17 - cons.price.idx: consumer price index - monthly indicator (numeric)
18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric)
19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
20 - nr.employed: number of employees - quarterly indicator (numeric)

#### Output variable (desired target):
21 - y - has the client subscribed a term deposit? (binary: 'yes','no')

## EDA

### Cleaning data, new variable generation, and/or handling missing logistics

##### Cleaning Data
 $ age           : int
 $ job           : Factor 
 $ marital       : Factor 
 $ education     : Factor 
 $ default       : Factor 
 $ housing       : Factor 
 $ loan          : Factor 
 $ contact       : Factor 
 $ month         : Factor 
 $ day_of_week   : Factor 
 $ duration      : int
 $ campaign      : int
 $ pdays         : int
 $ previous      : int
 $ poutcome      : Factor 
 $ emp.var.rate  : Factor 
 $ cons.price.idx: Factor 
 $ cons.conf.idx : Factor 
 $ euribor3m     : Factor 
 $ nr.employed   : Factor 
 $ y             : Factor 
```{r}
# read in the data
df <- read.csv2('../../Data/bank-additional-full.csv')

# get summary statistics
summary(df)

# types of each category
str(df)

```

#####New Variable Generation
```{r}

```

##### Handling of Missing Logistics
There are no missing values 
```{r}
# check for missing values
sapply(df, function(x) sum(is.na(x)))
```


2.  There are way more No's than Yes for the response variable.  Consider sampling your training data set in such a way that the Yes and No's are more equally balanced.   (This technique is referred to as downsampling.  There is also upsampling.)  After your groups meet and discuss, you can meet with me for a deeper discussion on this issue.
```{r}

```


3.  Of course EDA.  PCA_Unit9 R script has some potential starting graphics you can sift through as well as doing some PCA.  We'll discuss additional strategies in later units.  Don't forget summary statistics.  
```{r}

```