---
title: "Proj2_EDA_GraceLang"
author: "Grace Lang"
date: "7/11/2020"
output: word_document
---
# Data set
https://archive.ics.uci.edu/ml/datasets/bank%20marketing
# Paper Referenced - that uses the dataset and conclusions
http://media.salford-systems.com/video/tutorial/2015/targeted_marketing.pdf

## Response
21 - y - has the client subscribed a term deposit? (binary: 'yes','no')

# Questions about dataset
I'm not sure what the following variables mean in the context of the response:

11 - duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.

# other attributes:
12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
14 - previous: number of contacts performed before this campaign and for this client (numeric)
15 - poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')

# social and economic context attributes
16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
17 - cons.price.idx: consumer price index - monthly indicator (numeric)
18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric)
19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
20 - nr.employed: number of employees - quarterly indicator (numeric)



```{r setup, include=FALSE}
library(naniar)

#Grace's connection
bank <- read.csv2("C:/Users/david/OneDrive/Desktop/AppliedStats/Project2_Applied_Stats/bank-additional-full.csv")

summary(bank)
#pdays may be a irrelevant data point

#No variables missing?
gg_miss_var(bank)
```



# Having  issues plotting categorical - not sure what i'm doing wrong here
Need to find a better alternative for these views... what I iterated on wasn't great
```{r}
library(dplyr)

#Categorical x Categorical
plot(bank$y, bank$job, xlab = "Response", ylab = "job")

#Categorical x Continuous
plot(bank$y, bank$pdays, xlab = "Response", ylab = "pdays")
plot(bank$y, bank$previous, xlab = "Response", ylab = "previous")
plot(bank$y, bank$poutcome, xlab = "Response", ylab = "poutcome")
plot(bank$y, bank$emp.var.rate, xlab = "Response", ylab = "emp.var.rate")
plot(bank$y, bank$cons.price.idx, xlab = "Response", ylab = "cons.price.idx")
plot(bank$y, bank$cons.conf.idx, xlab = "Response", ylab = "cons.conf.idx")
plot(bank$y, bank$euribor3m, xlab = "Response", ylab = "euribor3m")
plot(bank$y, bank$nr.employed, xlab = "Response", ylab = "nr.employed")


```
