---
title: "Proj2_EDA_JuliaLayne"
author: "Julia Layne"
date: "7/12/2020"
output: word_document
---
# Data set
https://archive.ics.uci.edu/ml/datasets/bank%20marketing
# Paper Referenced - that uses the dataset and conclusions
http://media.salford-systems.com/video/tutorial/2015/targeted_marketing.pdf

## Response
21 - y - has the client subscribed a term deposit? (binary: 'yes','no')

# Questions about dataset
I'm not sure what the following variables mean in the context of the response:

11 - duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.

# other attributes:
12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
14 - previous: number of contacts performed before this campaign and for this client (numeric)
15 - poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')

# social and economic context attributes
16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
17 - cons.price.idx: consumer price index - monthly indicator (numeric)
18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric)
19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
20 - nr.employed: number of employees - quarterly indicator (numeric)



```{r}
library(naniar)
library(gplots)
library(ggplot2)
library(dplyr)

#Bank
bank <- read.csv2("bank-additional-full.csv",header = TRUE)
summary(bank)
#pdays may be a irrelevant data point
#Response is heavily weighted on No's 

# ?? WOULD WE EVEN NEED TO SCALE ANY OF THESE CATEGORICAL VARIABLES? Q for Turner

#No variables missing?
gg_miss_var(bank)
summary(bank)
bank
```

```{r}

#Clean up whitespaces

#Unknowns
clean_bank<-bank

#1 - age (numeric)
# okay
#2 - job : type of job (categorical: 'admin.','blue-collar','entrepreneur','housemaid','management','retired','self-employed','services','student','technician','unemployed','unknown')
clean_bank$job <- factor(clean_bank$job)
#3 - marital : marital status (categorical: 'divorced','married','single','unknown'; note: 'divorced' means divorced or widowed)
clean_bank$marital <- factor(clean_bank$marital)
#4 - education (categorical: 'basic.4y','basic.6y','basic.9y','high.school','illiterate','professional.course','university.degree','unknown')
clean_bank$education <- factor(clean_bank$education)
#5 - default: has credit in default? (categorical: 'no','yes','unknown')
clean_bank$default <- factor(clean_bank$default)
#6 - housing: has housing loan? (categorical: 'no','yes','unknown')
clean_bank$housing <- factor(clean_bank$housing)
#7 - loan: has personal loan? (categorical: 'no','yes','unknown')
clean_bank$loan <- factor(clean_bank$loan)


# related with the last contact of the current campaign:
#8 - contact: contact communication type (categorical: 'cellular','telephone') 
clean_bank$contact <- factor(clean_bank$contact)
#9 - month: last contact month of year (categorical: 'jan', 'feb', 'mar', ..., 'nov', 'dec')
clean_bank$month <- factor(clean_bank$month)
#10 - day_of_week: last contact day of the week (categorical: 'mon','tue','wed','thu','fri')
clean_bank$day_of_week <- factor(clean_bank$day_of_week)
#11 - duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.

# other attributes:
#12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
#13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
#14 - previous: number of contacts performed before this campaign and for this client (numeric)
#15 - poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')
# social and economic context attributes

#16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
clean_bank$int.emp.var.rate<-as.integer(clean_bank$emp.var.rate)
clean_bank$emp.var.rate <- factor(clean_bank$emp.var.rate)
#17 - cons.price.idx: consumer price index - monthly indicator (numeric) 
clean_bank$int.cons.price.idx<-as.integer(clean_bank$cons.price.idx)
clean_bank$cons.price.idx <- factor(clean_bank$cons.price.idx)
#18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric) 
clean_bank$int.cons.conf.idx<-as.integer(clean_bank$cons.conf.idx)
clean_bank$cons.conf.idx <- factor(clean_bank$cons.conf.idx)
#19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
clean_bank$int.euribor3m<-as.integer(clean_bank$euribor3m)
clean_bank$euribor3m <- factor(clean_bank$euribor3m)
#20 - nr.employed: number of employees - quarterly indicator (numeric)
clean_bank$int.nr.employed<-as.integer(clean_bank$nr.employed)
clean_bank$nr.employed <- factor(clean_bank$nr.employed)

#Output variable (desired target):
#21 - y - has the client subscribed a term deposit? (binary: 'yes','no')
clean_bank$y <- factor(clean_bank$y)

clean_bank
```

